---
title: "2_main_analyses"
author: "ASM"
date: "26/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(dplyr)
library(lme4)
library(lmerTest)
library(rstatix)
library(broom)
library(janitor)
library(glmmTMB)
library(DHARMa)
library(betareg)
library(emmeans)
library(ggeffects)
library(mvtnorm)
library(boot)
library(bbmle)
```


#----1. load clean data
```{r}

load(here("anonymized_data/final_merged_data.Rda"))

```


#----2. explore demographics
```{r}

n<- final_data%>%
  distinct(unique_id) #n of 552

gender<- final_data %>%
  distinct(unique_id, gender)%>%
  group_by(gender)%>%
  tally()
#16 na for gender I think from LEQs withouth associated demographics

range(final_data$age, na.rm = T) #4.33 - 30.9 months so 4 - 31 if we round.
```

#------ 4,make french exposure, english exposure and heritage language exposure variables

```{r}


##NOTE: Originally I had also added exposure on language 3 and 4. Language 3 and 4 always amount to less than 10% because we excluded trilinguals. However, 
#in parental interviews caregivers frequently give a third language and the final calculation is so low it amounts to 0 
#this was 0 inflating our english and french exposures, and because we are interested in how bilingual strategies relate to bilingual exposure, I decided to get rid of l3 and 4 exposure when making the frenc, english, and heritage language esposure variables.
            
  

vars<-c("exp_l3", "exp_l4")

final_data_language_coded<-
  final_data %>%
 mutate_at(vars, as.numeric) %>%
  mutate(l3 = replace(l3, l3=="NA", "no lang")) %>%
  mutate(l4 = replace(l4, l4=="NA", "no lang")) %>%
  mutate(fre_exp = case_when(l1=="French"~ exp_l1,
                                l2=="French"~ exp_l2,
                                
                             #l3=="French" & exp_l3 >=1 ~ exp_l3, 
                             #l4=="French"  & exp_l4 >=1 ~ exp_l4
                             ))%>%
    mutate(eng_exp = case_when(l1=="English"~ exp_l1,
                                l2=="English"~ exp_l2,
                                #l3=="English"~ exp_l3,
                                #l4=="English"~ exp_l4
                               )) %>%
  mutate(min_exp = case_when (!(l1 %in% c("French", "English"))& l2 %in% c("French", "English") ~ exp_l1,
                              !(l2 %in% c("French", "English")) & l1 %in% c("French", "English") ~ exp_l2
                              ))


```


#-----3. categorize LEQ data into family language strategies
```{r}
final_data_strat <- final_data_language_coded%>%
  mutate(strategy=case_when(
    
  #one-parent-one-lang-strict: Each parent speaks only one language and both parents don't speak the same language
    care1_l1=="regularly" & care1_l2 %in% c("never", "NA") & care2_l1 %in% c("never", "NA") & care2_l2=="regularly"~ "one-parent-one-lang",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1=="regularly" &care2_l2 %in% c("never", "NA")~ "one-parent-one-lang",
    
  #one-parent-one-lang-flex: Each parent speaks mainly one language, and one of them sometimes speaks another language
  #(The same language cannot be spoken regularly by both parents or that would be considered a one-parent-bilingual strategy)
    #**NOTE: we ended up having a single category for OPOL instead of divinding it into strict and flex.
    care1_l1=="regularly" & care1_l2 %in% c("never", "NA") & care2_l1=="sometimes" &care2_l2=="regularly"~ "one-parent-one-lang",
    care1_l1=="regularly" & care1_l2=="sometimes" & care2_l1 %in% c("never", "NA") &care2_l2=="regularly"~ "one-parent-one-lang",
    care1_l1=="sometimes" & care1_l2=="regularly" & care2_l1=="regularly" &care2_l2 %in% c("never", "NA")~ "one-parent-one-lang",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1=="regularly" &care2_l2=="sometimes"~ "one-parent-one-lang",
    
    
  #one-parent-bilingual: One parent speaks 2 languages, the other parent only speaks one language
    
    #Starting point
    care1_l1== "regularly" & care1_l2=="regularly" & care2_l1 == "regularly" & care2_l2 %in% c("never", "NA") ~ "one-parent-bilingual",
    care1_l1== "regularly" & care1_l2=="regularly" & care2_l1 %in% c("never", "NA") & care2_l2=="regularly" ~ "one-parent-bilingual",
    care1_l1== "regularly" & care1_l2 %in% c("never", "NA") & care2_l1 == "regularly" & care2_l2=="regularly" ~ "one-parent-bilingual",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1 == "regularly" & care2_l2=="regularly" ~ "one-parent-bilingual",

    #Some bilingual parents put both of their languages as sometimes rather than both as regularly
    care1_l1== "sometimes" & care1_l2=="sometimes" & care2_l1 == "regularly" & care2_l2 %in% c("never", "NA") ~ "one-parent-bilingual",
    care1_l1== "sometimes" & care1_l2=="sometimes" & care2_l1 %in% c("never", "NA") & care2_l2=="regularly" ~ "one-parent-bilingual",
    care1_l1== "regularly" & care1_l2 %in% c("never", "NA") & care2_l1 == "sometimes" & care2_l2=="sometimes" ~ "one-parent-bilingual",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1 == "sometimes" & care2_l2=="sometimes" ~ "one-parent-bilingual",

    #One parent bilingual (R&R or S&S) and another parent that speaks one of the language but not the other (R|S & N|NA)
    care1_l1 == "regularly" & care1_l2=="regularly" & care2_l1 %in% c("regularly", "sometimes") & care2_l2 %in% c("never", "NA") ~ "one-parent-bilingual",
    care1_l1 == "regularly" & care1_l2=="regularly" & care2_l1 %in% c("never", "NA") & care2_l2 %in% c("regularly", "sometimes") ~ "one-parent-bilingual",
    care1_l1 %in% c("regularly", "sometimes") & care1_l2 %in% c("never", "NA") & care2_l1 == "regularly" & care2_l2=="regularly" ~ "one-parent-bilingual",
    care1_l1 %in% c("never", "NA") & care1_l2 %in% c("regularly", "sometimes") & care2_l1 == "regularly" & care2_l2=="regularly" ~ "one-parent-bilingual",

    care1_l1 == "sometimes" & care1_l2=="sometimes" & care2_l1 %in% c("regularly", "sometimes") & care2_l2 %in% c("never", "NA") ~ "one-parent-bilingual",
    care1_l1 == "sometimes" & care1_l2=="sometimes" & care2_l1 %in% c("never", "NA") & care2_l2 %in% c("regularly", "sometimes") ~ "one-parent-bilingual",
    care1_l1 %in% c("regularly", "sometimes") & care1_l2 %in% c("never", "NA") & care2_l1 == "sometimes" & care2_l2=="sometimes" ~ "one-parent-bilingual",
    care1_l1 %in% c("never", "NA") & care1_l2 %in% c("regularly", "sometimes") & care2_l1 == "sometimes" & care2_l2=="sometimes" ~ "one-parent-bilingual",
    
    #Both parents speak the same language regularly, and one of them speak another language sometimes 
    #(not one-parent-one-lang-flex because both parents use the same language regularly in opposition to each parent having a different language that they use regularly)
    care1_l1=="regularly" & care1_l2=="sometimes" & care2_l1=="regularly" & care2_l2 %in% c("never", "NA") ~ "one-parent-bilingual",
    care1_l1=="regularly" & care1_l2 %in% c("never", "NA") & care2_l1=="regularly" & care2_l2=="sometimes" ~ "one-parent-bilingual",
    care1_l1=="sometimes" & care1_l2=="regularly" & care2_l1 %in% c("never", "NA")& care2_l2=="regularly" ~ "one-parent-bilingual",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1=="sometimes" & care2_l2=="regularly" ~ "one-parent-bilingual",
    
    
  #both-parents-bilingual: Both parents speak both languages either regularly or sometimes
    care1_l1%in% c("regularly", "sometimes") & care1_l2%in% c("regularly", "sometimes") & care2_l1%in% c("regularly", "sometimes") &care2_l2%in% c("regularly", "sometimes") ~ "both-parents-bilingual",

  #one-parent-one-lang-flex: Each parent speaks mainly one language, and one of them sometimes speaks another language
  #The same language cannot be spoken regularly by both parents or it comes to be filled under the one-parent-bilingual strategy.
  #**NOTE: we ended up having a single category for OPOL instead of divinding it into strict and flex.
    care1_l1=="regularly" & care1_l2 %in% c("never", "NA") & care2_l1=="sometimes" &care2_l2=="regularly"~ "one-parent-one-lang",
    care1_l1=="regularly" & care1_l2=="sometimes" & care2_l1 %in% c("never", "NA") &care2_l2=="regularly"~ "one-parent-one-lang",
    care1_l1=="sometimes" & care1_l2=="regularly" & care2_l1=="regularly" &care2_l2 %in% c("never", "NA")~ "one-parent-one-lang",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1=="regularly" &care2_l2=="sometimes"~ "one-parent-one-lang",
    
  #one-language-at-home: There is only one language spoken in the home by both parents
    care1_l1=="regularly" & care1_l2 %in% c("never", "NA") & care2_l1=="regularly" &care2_l2 %in% c("never", "NA")~ "one-language-at-home",
    care1_l1 %in% c("never", "NA") & care1_l2=="regularly" & care2_l1 %in% c("never", "NA") & care2_l2=="regularly"~ "one-language-at-home",
    
    care1_l1=="sometimes" & care1_l2 %in% c("never", "NA") & care2_l1=="sometimes" &care2_l2 %in% c("never", "NA")~ "one-language-at-home",
    care1_l1 %in% c("never", "NA") & care1_l2=="sometimes" & care2_l1 %in% c("never", "NA") & care2_l2=="sometimes"~ "one-language-at-home",
    
  #single-parent: Family in which there is only one caregiver
    care1_l1 %in% c("never", "NA") & care1_l2 %in% c("never", "NA") & care1_l3 %in% c("never", "NA") & care1_l4 %in% c("never", "NA") ~ "single-parent",
    care2_l1 %in% c("never", "NA") & care2_l2 %in% c("never", "NA") & care2_l3 %in% c("never", "NA") & care2_l4 %in% c("never", "NA") ~ "single-parent"
  
    ))
    

#Calculate the proportion for the different strategies
prop_strategy<-final_data_strat%>%
group_by(strategy) %>%
  tally() %>%
  mutate(prop = n/sum(n))



# I decided to exclude "true" monolinguals characterized here to the best of our abilities as children whose families use one language at home strategy, and the childcare language is the same as the one language at home strategy = they are being exposed to a single language = most likely not a bilingual strategy.

final_data_strat <-
  final_data_strat %>%
  mutate(bilinugal_strategy= case_when(strategy=="one-language-at-home" & childcare_l1 == l1 & childcare_l2 %in% c(NA, "NA")  ~ "no",
                                       strategy=="one-language-at-home" & childcare_l2 == l1 & childcare_l1 %in% c(NA, "NA") ~"no",
                                       strategy=="one-language-at-home" & childcare_l1 %in% c(NA, "NA") & childcare_l2 %in% c(NA, "NA") ~"no",
                                       TRUE~"yes")) %>%
  filter(bilinugal_strategy=="yes")


save(final_data_strat, file=here("anonymized_data/final_data_strat.Rda"))

#Sanity check, all the monolinguals we have are using bilingual strategies 
eng_monoling <- final_data_strat %>%
  filter(eng_exp == 100)

fre_moling <- final_data_strat %>%
  filter(fre_exp==100)
```

#---- coding minority versus majority language
```{r}

final_data_strat <- final_data_strat %>% 
  mutate(lang_status = case_when(
    l2=="NA" ~ "monolingual",
    exp_l2== 0 ~ "monolingual",
    l1 %in% c("French", "English") & l2 %in% c("French", "English") ~ "both_maj",
     l1 %in% c("French", "English") & !(l2 %in% c("French", "English")) ~ "one_maj",
    !(l1 %in% c("French", "English")) & l2 %in% c("French", "English") ~ "one_maj",
    TRUE ~ "both_min"
  ))

#do check that the monolinguals are actually monolinguals and not both min


#subseting to get data only from each child's first visit to the lab. For main analyses
first_visit_only <- final_data_strat%>%
    arrange(anon_baby_id, age) %>%
  group_by(anon_baby_id) %>%
  mutate(visit = row_number()) %>%
  ungroup() %>%
  filter (visit==1)

```


###PERFORMING ANOVAS
```{r}
#filtering the single parent cases because there are not enough to have any power statistically.
for_anovas<- first_visit_only %>%
  filter(!strategy == "single-parent")

anova_model_french <- aov(fre_exp ~ strategy, data= for_anovas)

tidy(anova_model_french)
tukey_hsd(anova_model_french)


anova_model_english <- aov(eng_exp ~ strategy, data= for_anovas)
tidy(anova_model_english)
tukey_hsd(anova_model_english)

anova_model_heritage <- aov(min_exp ~ strategy, data= for_anovas)
tidy(anova_model_heritage)
```


```{r}
for_t_tests<- final_data_strat %>%
  select(c(strategy, lang_status, fre_exp, eng_exp, min_exp)) 

pwc_exp_fre <- for_t_tests %>% 
  pairwise_t_test(
    fre_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    ) 

pwc_exp_eng <- for_t_tests %>% 
  pairwise_t_test(
    eng_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    ) 

pwc_exp_min <- for_t_tests %>% 
  pairwise_t_test(
    min_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    ) 

save(pwc_exp_fre, file=here("results/t_fre.Rda"))
save(pwc_exp_eng, file=here("results/t_eng.Rda"))
save(pwc_exp_min, file=here("results/t_min.Rda"))
```


#####----------------- BY GROUP ANALYSES--------------------####
#------ Group 1: When both l1 and l2 are majority languages

```{r}

both_maj_data<- final_data_strat %>%
  filter(lang_status == "both_maj")
#355 observations


#Calculate the proportion for the different strategies in this group
prop_strategy_group1<-both_maj_data %>%
group_by(strategy) %>%
  tally() %>%
  mutate(prop = n/sum(n))

print(prop_strategy_group1)

for_t_tests_g1<- both_maj_data %>%
  select(c(strategy, lang_status, fre_exp, eng_exp))

#french exposure
pwc_fre_exp_g1 <- for_t_tests_g1 %>% 
  pairwise_t_test(
    fre_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_fre_exp_g1

#english exposure
pwc_eng_exp_g1 <- for_t_tests_g1 %>% 
  pairwise_t_test(
    eng_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_eng_exp_g1


```


#------ Group 2: When only l1 or 2 are a majority language, and the other pair is a minority lang

```{r}
one_maj_data<- final_data_strat %>%
  filter(lang_status == "one_maj")

#80 observations

prop_strategy_group2<-one_maj_data %>%
group_by(strategy) %>%
  tally() %>%
  mutate(prop = n/sum(n))

print(prop_strategy_group2)

#the cell sizes are a bit small... but it does give a different story about which strategies are the most used than when both languages are majority

#get the minority language exposure
one_maj_data<- one_maj_data %>%
  mutate(min_exp = case_when (!(l1 %in% c("French", "English"))& l2 %in% c("French", "English") ~ exp_l1,
                              !(l2 %in% c("French", "English")) & l1 %in% c("French", "English") ~ exp_l2,
                              TRUE ~ 0
                              ))

for_t_tests_g2<- one_maj_data %>%
  select(c(strategy, lang_status, fre_exp, eng_exp, min_exp))

#french exposure
pwc_fre_exp_g2 <- for_t_tests_g2 %>% 
  pairwise_t_test(
    fre_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_fre_exp_g2

#english exposure
pwc_eng_exp_g2 <- for_t_tests_g2 %>% 
  pairwise_t_test(
    eng_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_eng_exp_g2

#minority exposure
pwc_min_exp_g2 <- for_t_tests_g2 %>% 
  pairwise_t_test(
    min_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_min_exp_g2


```

#--------- linear mixed models
```{r}

m1 <- lmer(exp_l1 ~ strategy + (1|anon_baby_id), data=final_data_strat) #can't with unique id, just anon. To check?

summary(m1)
#some significant effects by strategy.
#quite a bit of variance from the by-subject random intercept = balacedeness has to do also with individual variability than langauge strategy.
#take out the single parent language
#Maybe collapse the OPOL groups?

m1.1 <- lmer(exp_l1 ~ strategy + lang_status + (1|anon_baby_id), data=final_data_strat) 
summary(m1.1)
#L1 is many language code everything relative to French as the official language. 

m1.2 <- lmer(exp_l1 ~ strategy * lang_status + (1|anon_baby_id), data=final_data_strat) 
summary(m1.2)


anova(m1, m1.1) #significantly different so langauge status matters but is it due to the monolinguals?

test<- lmer(exp_l1 ~ lang_status + (1|anon_baby_id), data=final_data_strat) 
summary(test) #maybe monolinguals are driving the effect. Which makes sense because ofc monolinguals are unbalanced.


#prediction of French exposure
m2 <- lmer(fre_exp ~ strategy + (1|anon_baby_id), data=final_data_strat) 
m2.2<- lmer(fre_exp ~ strategy * lang_status + (1|anon_baby_id), data=final_data_strat) 
summary(m2.2)
summary(m2)
#Both parents bilingual and opol-strict marginally

#Look at heritage language speakers and look at the french exposure. 
#visualization more similar to analyses. Mean l1 exposure by strategy l1 exp in y.
#Do we just want to do t-tests instead .. order the strategies by how much l1 exposure they give you. And then heritage vs non heritage.

```


#T-test pairwise comparisons
```{r}
for_t_tests<- final_data_strat %>%
  select(c(strategy, lang_status, fre_exp, exp_l1)) 

#Order l1 amount of exposure by strategy
l1_by_strategy <- for_t_tests %>%
  select(c(strategy, exp_l1)) %>%
  group_by(strategy) %>%
  summarise(mean(exp_l1))

#French exposure
fre_by_strategy <- for_t_tests %>%
  select(c(strategy, fre_exp)) %>%
  group_by(strategy) %>%
  summarise(mean(fre_exp))

#Pairwise t test comparisons for all the strategies against amount of l1 exposure. Bonferroni corrected p for multiple comparisons. 
pwc_exp_l1 <- for_t_tests %>% 
  pairwise_t_test(
    exp_l1~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_exp_l1

#Pairwise t test comparisons for all the strategies against amount of french exposure. Bonferroni corrected p for multiple comparisons.
pwc_fre_exp <- for_t_tests %>% 
  pairwise_t_test(
    fre_exp~strategy, pool.sd = FALSE,
    p.adjust.method = "bonferroni"
    )
pwc_fre_exp




#langauge status
t.test(fre_exp~lang_status, data = for_t_tests, subset = lang_status %in% c("maj","min"),  mu=0, conf.level = 0.95, paired=F) #If you speak a majority langauge you are more likely to have more fre exp

#non Heritage englissh and french, vs.heritage. 


t.test(exp_l1~lang_status, data = for_t_tests, subset = lang_status %in% c("maj","min"),  mu=0, conf.level = 0.95, paired=F) #weather the langauges you speak at home are majority or minority does not influence the amount of L1 you get


```


#---------- create and save longitudinal data frame
```{r}
#keeps only data where the baby participated more than once
longitudinal_data<- final_data_strat %>%
  add_count(anon_baby_id) %>%
  filter(nn>1)

save(longitudinal_data, file=here("anonymized_data/longitudinal_data.Rda"))
```



#Explore beta regressions
```{r}
# Create the dataframes to run the models for each language. Monolinguals are excluded, as are the rows 
# where exposure for a given language equals 100. Since exposure is on a 0-100 scale and the beta distributions 
# works with values ranging from 0-1, exposure is divided by 100. 
for_beta <- for_anovas %>% 
  select(gender, anon_baby_id, age, fre_exp, eng_exp, min_exp, strategy, lang_status) %>% 
  filter(lang_status != "monolingual") %>% 
  filter(gender != "NA")

# Dataframe for english exposure
for_beta_eng <- for_beta %>% 
  select(age, eng_exp, strategy, lang_status, anon_baby_id) %>% 
  filter(eng_exp != "NA") %>% 
  filter(eng_exp !=100) %>% 
  mutate(eng_exp = eng_exp/100,
         strategy = as.factor(strategy), 
         lang_status = as.factor(lang_status))

# Dataframe for french exposure
for_beta_fre <- for_beta %>% 
  select(age, fre_exp, strategy, lang_status, anon_baby_id) %>%
  filter(fre_exp != "NA") %>% 
  filter(fre_exp != 100) %>% 
  mutate(fre_exp = fre_exp/100, 
         strategy = as.factor(strategy), 
         lang_status = as.factor(lang_status))

# Dataframe for minority language exposure
for_beta_min <- for_beta %>% 
  select(age, min_exp, strategy, lang_status, anon_baby_id) %>%
  filter(min_exp != "NA") %>% 
  filter(min_exp !=100) %>% 
  mutate(min_exp = min_exp/100,
         strategy = as.factor(strategy), 
         lang_status = as.factor(lang_status))


# When modelling direct proportions, using a beta distribution is a good idea. 
# We can fit a beta model with eng_exp and age as a predictor. 
# These models are better than the linear models, but still present issues (i.e., heteroscedasticity).

m2a <- glmmTMB(eng_exp ~ strategy, data=for_beta_eng, family= beta_family(link="logit")) 
summary(m2a)
sim_m2a <- simulateResiduals(m2a)
plot(sim_m2a)

m2b <- glmmTMB(eng_exp ~ strategy * age, data=for_beta_eng, family= beta_family(link="logit"))
summary(m2b) # NOTE: No significant effects
sim_m2b <- simulateResiduals(m2b)
plot(sim_m2b)

# To solve the heteroscedasticity problem, we can use the dispformula argument so that that the dispersion parameter phi is allowed to vary by the levels of "strategy"
m3a <- glmmTMB(eng_exp ~ strategy, data=for_beta_eng, family= beta_family(link="logit"), dispformula = ~strategy)
summary(m3a) # Significant effect for one-parent-bilingual; AIC = -83.7
sim_m3a <- simulateResiduals(m3a)
plot(sim_m3a) # MODEL OK!

m3b <- glmmTMB(eng_exp ~ strategy * age, data=for_beta_eng, family= beta_family(link="logit"), dispformula = ~strategy)
summary(m3b) # No significant effects
sim_m2b <- simulateResiduals(m2b)
plot(sim_m2b) # NOTE: Quantile deviations detected

# We can enhance the model by allowing the dispersion parameter in the beta distribution to differ across the different levels of "strategy" and the different values of "age"
m4a <- glmmTMB(eng_exp ~ strategy, data=for_beta_eng, family= beta_family(link="logit"), dispformula = ~ strategy * age)
summary(m4a)
sim_m4a <- simulateResiduals(m2b)
plot(sim_m4a) # NOTE: Quantile deviations detected

m4b <- glmmTMB(eng_exp ~ strategy * age, data=for_beta_eng, family= beta_family(link="logit"), dispformula = ~ strategy * age)
summary(m4b) # Significant effects for one-language-at-home and its interaction with age

# We run model diagnostics with DHARMa
sim_m4b <- simulateResiduals(m4b)
plot(sim_m4b) # MODEL OK!
testDispersion(sim_m4b) # MODEL OK!

# We can plot the fixed effects for m4b, the best model so far (very interesting!)
m4b_effects <- ggpredict(m4b, terms = c("age", "strategy"))
custom_palette = c("#BD081C","#0099FF", "#00B489", "#FF8F00")
plot(m4b_effects, add.data = F, colors= custom_palette, alpha=.2)


# We can estimate and compare the marginal means of the response variable across different levels of "strategy", while accounting for the effect of "age" and the varying dispersion parameter. This code computes pairwise comparisons of the estimated marginal means of eng_exp across the different levels of "strategy", while holding "age" constant at its mean value.
emtrends(m4b, pairwise ~ strategy, var="age") # trend for one-language-at-home significantly different from that of one parent-bilingual


# MODEL FOR FRENCH EXPOSURE + VISUALIZATION
m5 <- glmmTMB(fre_exp ~ strategy * age, data=for_beta_fre, family= beta_family(link="logit"), dispformula = ~ strategy * age)
summary(m5) # No significant results

# We run model diagnostics with DHARMa
sim_m5 <- simulateResiduals(m5)
plot(sim_m5) # MODEL OK!
testDispersion(sim_m5) # MODEL OK!

# We can plot the fixed effects for m5, the best model so far
m5_effects <- ggpredict(m5, terms = c("age", "strategy"))
custom_palette = c("#BD081C","#0099FF", "#00B489", "#FF8F00")
plot(m5_effects, add.data = F, colors= custom_palette, alpha=.2)

# We can estimate and compare the marginal means of the response variable across different levels of "strategy", while accounting for the effect of "age" and the varying dispersion parameter. 
emtrends(m5, pairwise ~ strategy, var="age") # no significant results

# MODEL FOR MINORITY LANGUAGE EXPOSURE + VISUALIZATION
m6 <- glmmTMB(min_exp ~ strategy * age, data=for_beta_min, family= beta_family(link="logit"), dispformula = ~ strategy + age)
summary(m6) # Significant effects for one-language-at-home and its interaction with age and the interaction between one-parent-one-lang and age.

# We run model diagnostics with DHARMa
sim_m6 <- simulateResiduals(m6)
plot(sim_m6) # MODEL OK!
testDispersion(sim_m6) # MODEL OK!

# We can plot the fixed effects for m6, the best model so far (very interesting!)
m6_effects <- ggpredict(m6, terms = c("age", "strategy"))
custom_palette = c("#BD081C","#0099FF", "#00B489", "#FF8F00")
plot(m6_effects, add.data = F, colors= custom_palette, alpha=.2)

# We can estimate and compare the marginal means of the response variable across different levels of "strategy", while accounting for the effect of "age" and the varying dispersion parameter. 
emtrends(m6, pairwise ~ strategy, var="age") # # trend for one-language-at-home significantly different from that of one parent-bilingual

```

